---
title: "HW4Code"
format: html
editor: visual
editor_options: 
  chunk_output_type: inline
  comments:
  utterances:
    repo: https://github.com/Nick-M52/Homework4.git
---



1. How does fish length predict fish weightfor trout perch (across all sample years)?

Problem 1-

1) Null Hypothesis: There is no statistically significant  relationship between the lengths of fish and their weight for trout perch.

Alternative Hypothesis:There is a statistically significant relationship between the lengths of fish and their weight for trout perch.


This is the code to clean up the data provided by the data creators:

```{r}

# Package ID: knb-lter-ntl.6.34 Cataloging System:https://pasta.edirepository.org.
# Data set title: North Temperate Lakes LTER: Fish Lengths and Weights 1981 - current.
# Data set creator:  John Magnuson - University of Wisconsin-Madison 
# Data set creator:  Stephen Carpenter - University of Wisconsin-Madison 
# Data set creator:  Emily Stanley - University of Wisconsin-Madison 
# Metadata Provider:  NTL Information Manager - University of Wisconsin-Madison 
# Contact:    -  NTL LTER  - ntl.infomgr@gmail.com
# Stylesheet v2.11 for metadata conversion into program: John H. Porter, Univ. Virginia, jporter@virginia.edu 

inUrl1  <- "https://pasta.lternet.edu/package/data/eml/knb-lter-ntl/6/34/7dddb5e70250d2e24537e34272316220" 
infile1 <- tempfile()
try(download.file(inUrl1,infile1,method="curl"))
if (is.na(file.size(infile1))) download.file(inUrl1,infile1,method="auto")

                   
 dt1 <-read.csv(infile1,header=F 
          ,skip=1
            ,sep=","  
                ,quot='"' 
        , col.names=c(
                    "lakeid",     
                    "year4",     
                    "sampledate",     
                    "gearid",     
                    "spname",     
                    "sampletype",     
                    "depth",     
                    "rep",     
                    "indid",     
                    "length",     
                    "weight",     
                    "sex",     
                    "fishpart",     
                    "spseq",     
                    "flag"    ), check.names=TRUE)
               
unlink(infile1)
		    
# Fix any interval or ratio columns mistakenly read in as nominal and nominal columns read as numeric or dates read as strings
                
if (class(dt1$lakeid)!="factor") dt1$lakeid<- as.factor(dt1$lakeid)
if (class(dt1$year4)=="factor") dt1$year4 <-as.numeric(levels(dt1$year4))[as.integer(dt1$year4) ]               
if (class(dt1$year4)=="character") dt1$year4 <-as.numeric(dt1$year4)                                   
# attempting to convert dt1$sampledate dateTime string to R date structure (date or POSIXct)                                
tmpDateFormat<-"%Y-%m-%d"
tmp1sampledate<-as.Date(dt1$sampledate,format=tmpDateFormat)
# Keep the new dates only if they all converted correctly
if(length(tmp1sampledate) == length(tmp1sampledate[!is.na(tmp1sampledate)])){dt1$sampledate <- tmp1sampledate } else {print("Date conversion failed for dt1$sampledate. Please inspect the data and do the date conversion yourself.")}                                                                    
rm(tmpDateFormat,tmp1sampledate) 
if (class(dt1$gearid)!="factor") dt1$gearid<- as.factor(dt1$gearid)
if (class(dt1$spname)!="factor") dt1$spname<- as.factor(dt1$spname)
if (class(dt1$sampletype)!="factor") dt1$sampletype<- as.factor(dt1$sampletype)
if (class(dt1$depth)=="factor") dt1$depth <-as.numeric(levels(dt1$depth))[as.integer(dt1$depth) ]               
if (class(dt1$depth)=="character") dt1$depth <-as.numeric(dt1$depth)
if (class(dt1$rep)=="factor") dt1$rep <-as.numeric(levels(dt1$rep))[as.integer(dt1$rep) ]               
if (class(dt1$rep)=="character") dt1$rep <-as.numeric(dt1$rep)
if (class(dt1$indid)!="factor") dt1$indid<- as.factor(dt1$indid)
if (class(dt1$length)=="factor") dt1$length <-as.numeric(levels(dt1$length))[as.integer(dt1$length) ]               
if (class(dt1$length)=="character") dt1$length <-as.numeric(dt1$length)
if (class(dt1$weight)=="factor") dt1$weight <-as.numeric(levels(dt1$weight))[as.integer(dt1$weight) ]               
if (class(dt1$weight)=="character") dt1$weight <-as.numeric(dt1$weight)
if (class(dt1$sex)!="factor") dt1$sex<- as.factor(dt1$sex)
if (class(dt1$fishpart)!="factor") dt1$fishpart<- as.factor(dt1$fishpart)
if (class(dt1$spseq)!="factor") dt1$spseq<- as.factor(dt1$spseq)
if (class(dt1$flag)!="factor") dt1$flag<- as.factor(dt1$flag)
                
# Convert Missing Values to NA for non-dates
                


# Here is the structure of the input data frame:
str(dt1)                            
attach(dt1)                            
# The analyses below are basic descriptions of the variables. After testing, they should be replaced.                 

summary(lakeid)
summary(year4)
summary(sampledate)
summary(gearid)
summary(spname)
summary(sampletype)
summary(depth)
summary(rep)
summary(indid)
summary(length)
summary(weight)
summary(sex)
summary(fishpart)
summary(spseq)
summary(flag) 
                # Get more details on character variables
                 
summary(as.factor(dt1$lakeid)) 
summary(as.factor(dt1$gearid)) 
summary(as.factor(dt1$spname)) 
summary(as.factor(dt1$sampletype)) 
summary(as.factor(dt1$indid)) 
summary(as.factor(dt1$sex)) 
summary(as.factor(dt1$fishpart)) 
summary(as.factor(dt1$spseq)) 
summary(as.factor(dt1$flag))
detach(dt1)               
        
fishlengths<-dt1

```

```{r}
head(fishlengths)

```

```{r}
#installing packages
library(tidyverse) # general usage
library(here) # organization
library(plotly) # interactive plots
library(skimr) # quick glimpses at data
library(naniar) # visualizing missing data
library(magick) # inserting images into plots
library(NatParksPalettes) # one example of color palette package
library(wesanderson) # another example of a color palette package
library(patchwork) # putting plots together
```

```{r}
library(performance)
library(broom)
library(flextable)
library(ggeffects)
library(car)
```


#2 code)

```{r}
#creating filtered data frame
library(dplyr)

# Specify the columns you want to keep
length_wt_yr <- c("year4", "length", "weight")

# Filter the data table to include only the specified columns
filtered_data <- fishlengths %>%
  select(all_of(length_wt_yr))

view(filtered_data)
```

```{r}
#visuaization of the missing data for the filtered dataset
gg_miss_var(filtered_data)

```
#2 answer)
a) The missing data is relevant to our hypothesis because there are several missing data points for weight which could alter the results. However, the 500+ data points we can reference should be good enough to test our hypothesis.

#3 code)
```{r}
#removing rows with NA data
complete_data<- filtered_data[complete.cases(filtered_data),]
view(complete_data)
```

```{r}
#exploring the relationship visually
ggplot(data = complete_data, aes(x = length, y = weight)) +
  geom_point()
```


```{r}
#linear regression test
lmlength_mass<- lm(weight~length, data= complete_data)
#show results
summary(lmlength_mass)
```
#4 code)
```{r}
#creating diagnostic plots to visually check assumptions
par(mfrow = c(2,2))
par(mar = c(2, 4, 2, 2))
plot(lmlength_mass)
```

#5 answer)

RvF- The residuals are not evenly and randomly distributed, so the assumption of homoscedasticity is violated.

S-L: These residuals are not evenly and randomly distributed as well, so the assumption of homoscedasticity is violated.

Q-Q: These residuals are not normally distribute because they do not follow the straight trend line.

RvL- These residuals show that outliers largely affect the data and that the assumption of homoscedasticity is violated.

We can infer that the residuals are not linear and that the data is heteroscedastic so the linear regression is not a good predictor of values.

#6 code)

```{r}
#summary of linear regression
summary(lmlength_mass)
#this summary shows a strong relationship between predictor and response. high r squared indicates good fit. intercept and length coefficients statistically significant
```
```{r}
#creating table to summarize anova table
model_summary<-summary(lmlength_mass)
#storing anova table as object
model_squares<- anova(lmlength_mass)
model_squares
```


```{r}
model_squares_table<-tidy(model_squares) %>% 
  # round the sum of squares and mean squares columns 
  mutate(across(sumsq:meansq, ~ round(.x, digits = 5))) %>% 
  # round the F-statistic 
  mutate(statistic = round(statistic, digits = 1)) %>% 
  # replace the small p value with < 0.001
  mutate(p.value = case_when(
    p.value < 0.001 ~ "< 0.001"
  )) %>% 
  # rename length to be more meaningful
  mutate(term = case_when(
    term == "length" ~ "Length (mm)",
    term == "weight" ~ "Weight (g)",
    TRUE ~ term
  )) %>% 
  # make the data frame a flextable object
  flextable() %>% 
  # change the header labels to be meaningful
  set_header_labels(df = "Degrees of Freedom", 
                    sumsq = "Sum of squares",
                    meansq = "Mean squares",
                    statistic = "F-statistic",
                    p.value = "p-value")

model_squares_table
```

#8 answer)

The information of the anova table suggests similar information to the summary table. It shows a strong correlation between the predictor and response, however the low p-value suggests that the null hypothesis should be rejected

#9 answer)

The linear regression revealed a significant direct relationship between the weight of the plants and their length (coefficient estimate= 3.00965, t(91088) = 369.7, p < 2e-16). The model accounted for a substantial amount of the variability in weight (R² = 0.6001), which was supported by the highly significant F-statistic (F(1, 91088) = 1.367e+05. These findings suggest that length is a strong predictor of weight, but there is a high variance in the data.



#10 code)

```{r}
# extract model predictions using ggpredict
predictions <- ggpredict(lmlength_mass, terms = "length")

predictions
```

```{r}

plot_predictions <- ggplot(data = complete_data, 
                           aes(x = length, y = weight)) +
  # first plot the underlying data from maples_data
  geom_point() +
  # then plot the predictions
  geom_line(data = predictions, 
            aes(x = x, y = predicted), 
            color = "blue", linewidth = 1) +
  # then plot the 95% confidence interval from ggpredict
  geom_ribbon(data = predictions, 
              aes(x = x, y = predicted, ymin = -312, ymax = 3621), 
              alpha = 0.1) +
  # theme and meaningful labels
  theme_bw() +
  labs(x = "Length (mm)",
       y = "Weight (g)")

plot_predictions
#the CI was not showing on the graph so I manually input the data which does not display them quite correctly but gives an idea of the CI
```
a) This plot shows the predicted weight value based on the length value with the blue line. The filtered data is plotted on this to show how it compares to the predicted values. The grey area is the confidence interval and the mean of the values is expected to fall within this.
